% MiniSys
% Copyright (C) 2015 Lumin Zhou <cdluminate@gmail.com>

\documentclass[10pt,a4paper]{article}
% use Cantarell (Gnome) as default English font
\usepackage[default]{cantarell}
\usepackage{framed}
\usepackage{geometry}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue]{hyperref}
\usepackage{fancyhdr}
%\usepackage{ctex}
\usepackage{listings}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{textcomp}

% configure
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\fancyhead[RO,LE]{Cook a Mini Bootable Linux System}

% title
\begin{titlepage}
\title{\huge\textbf{Cook a Mini Bootable Linux System}\\ \small{--- Grub2 + Kernel + Busybox ---}}
\author{C.D.Luminate\\ \small cdluminate@gmail.com}
\date{\today}
\end{titlepage}

% ----start document----
\begin{document}
\pagestyle{fancy}
% generate the title page
\maketitle
\hrulefill
% Generate a table of contents, namely index.
\tableofcontents

\newpage
\section{Introduction}
\subsection{Tools in Need}
Before you go ahead, you should check if you had these tools
installed in your system\footnote{we assume your machine is running Debian or its variant.}
\begin{enumerate}
\item build-essential, Depends: \\
	libc6-dev, gcc, g++, make, dpkg-dev
\item grub-common, grub2-common, grub-pc, grub-pc-bin
\item QEMU - fast processor emulator
\end{enumerate}

\section{Trace the Boot Process}
\input{trace.tex}

\section{Build Linux Kernel Image}
\subsection{Download kernel source}
Pick a kernel from the linux kernel archives.\cite{bib:kernel.org}\newline
Here I use the Debian redistributed one or linux 4.0 :
\begin{verbatim}
linux-3.16.7-ckt7-1
linux-4.0
\end{verbatim}
Then extract it to the workplace :
\begin{framed}\begin{verbatim}
$ tar zxvf linux-3.16.7-ckt7.tar.gz -C workplace/
$ tar zxvf linux-4.0.tar.gz -C workplace/
$ cd workplace
\end{verbatim}\end{framed}
\subsection{Configure the Kernel}
To simplify the Procedure, I just used the default kernel config for AMD64 architecture, so type
\begin{framed}\begin{verbatim}
$ cd workplace/linux-?/
$ make x86_64_defconfig
$ make menuconfig
\end{verbatim}\end{framed}
Modify some configurations as you like, via menuconfig.\footnote{For detail please look up other materials.}
\subsection{Compile kernel}
Lets compile the kernel. Maybe you should invoke "make help" at first.
\begin{framed}\begin{verbatim}
$ make -j4 vmlinux
$ make -j4 bzImage
\end{verbatim}\end{framed}
The process takes a long while.
\subsection{The kernel}
After compiling, the file \texttt{"arch/x86/boot/bzImage"} is exactly what we need.
\begin{framed}\begin{verbatim}
bzImage: Linux kernel x86 boot executable bzImage,
    version 3.16.7-ckt7 (lumin@debian) #2 SMP Sat Mar 21 09:15:07 UTC 2015, 
    RO-rootFS, swap_dev 0x5, Normal VGA
\end{verbatim}\end{framed}
Put this kernel file at proper place.

\section{Build Static Busybox}
\subsection{Download Busybox source}
You can download busybox source on official site.\cite{bib:busybox.net}\newline
Here I use Debian Redistributed one or another official one:
\begin{verbatim}
busybox-1.22.0-9+deb8u1
busybox-1.23.2.tar.bz2
\end{verbatim}
Extract the source pack and change directory into source tree.
\subsection{Configure Busybox}
\begin{framed}\begin{verbatim}
$ cd busybox-?/
$ make defconfig
$ make menuconfig
\end{verbatim}\end{framed}
Set the \texttt{"CONFIG\_STATIC=y"}, namely mark\newline
Busybox Settings - Build Options - ... Static Binary\newline
You can also mark the "dpkg" or something else as you like.

\subsection{Compile Busybox}
\begin{framed}\begin{verbatim}
$ make -j4 busybox
$ make install
\end{verbatim}\end{framed}
Then you will see a fine rootfs under directory \texttt{"\_install/"} .\newline
Copy all the content of \_install/ to workplace/initrd/:
\begin{framed}\begin{verbatim}
# cd _install
# mkdir -p workplace/initrd
# cp -av . workplace/initrd/
# chown -R root:root workplace/initrd/
\end{verbatim}\end{framed}
Now, Busybox preparation is completed. Lets Configure the system.

\section{Build Initramfs}
\textbf{HINT}: In Debian t/e \texttt{initramfs.img} is named \texttt{initrd.img} too.

\subsection{Make FHS available}
\begin{framed}\begin{verbatim}
# cd workplace/initrd/
# mkdir  boot bin dev proc sbin tmp boot etc 
# mkdir  lib root run srv usr home mnt sys var 
\end{verbatim}\end{framed}

\subsection{Configure initramfs files}
You can refer to the Debian package \texttt{base-files}.
\subsubsection{etc/fstab}
fstab stores static information about the filesystem, so let's vim etc/fstab.
\begin{framed}\begin{verbatim}
proc  /proc proc rw,nosuid,nodev,noexec,relatime 0 0
sysfs /sys sysfs rw,nosuid,nodev,noexec,relatime 0 0
tmpfs /run tmpfs rw,nosuid,relatime,mode=755     0 0
\end{verbatim}\end{framed}
Above are important items. If you would like to invoke
\begin{verbatim}
# mount -a
\end{verbatim}
in any script (like rcS or initramfs init) or manually, you should have this file.

\subsubsection{/dev/*}
\begin{framed}\begin{verbatim}
# cd wordplace/initrd
(# mknod -m 640 dev/initrd  b 1 250)
# mknod -m 600 dev/console c 5 1
# mknod -m 666 dev/null    c 1 3
\end{verbatim}\end{framed}

\subsubsection{etc/hostname}
Anything you like, such as \texttt{debian}.

\subsubsection{etc/hosts}
This is for basic network function.
\begin{framed}\begin{verbatim}
127.0.0.1  localhost debian
::1        localhost ip6-localhost ip6-loopback debian
ff02::1    ip6-allnodes
ff02::2    ip6-allrouters
\end{verbatim}\end{framed}

\subsubsection{/etc/inputrc}
Add this file to enable convenient keys.
\begin{framed}\begin{verbatim}
# /etc/inputrc
set input-meta on
set output-meta on
set bell-style none
$if mode=emacs
"\e[1~": beginning-of-line
"\e[4~": end-of-line
"\e[3~": delete-char
"\e[2~": quoted-insert
"\e[1;5C": forward-word
"\e[1;5D": backward-word
"\e[5C": forward-word
"\e[5D": backward-word
"\e\e[C": forward-word
"\e\e[D": backward-word
$if term=rxvt
"\e[8~": end-of-line
"\eOc": forward-word
"\eOd": backward-word
$endif
$endif
\end{verbatim}\end{framed}

\subsubsection{/etc/\{passwd,shadow\}}
this is passwd
\begin{verbatim}
root:x:0:0:root:/root:/bin/bash
\end{verbatim}
and this is shadow
\begin{verbatim}
root:::0:99999:7:::
\end{verbatim}
They are to enable root login and set root password as null.

\subsection{Initramfs init}
\textbf{Warning} : If there is no \texttt{/init} in initrd.img,
	kernel would regard the initrd.img as malformed/illegal one and then \textbf{panic}.
\begin{itemize}
\item If you don't want to create an \texttt{init} script, you can just link the init to busybox as following.
\item If you want to use a true init script, following is a very simple one that works.
\item Or even, you can write your own C init program.
\end{itemize}
Linking busybox init:
\begin{framed}\begin{verbatim}
# cd workplace/initrd/
# ln -s linuxrc init
\end{verbatim}\end{framed}
Creating a simple initramfs init script:
\begin{framed}\begin{verbatim}
#!/bin/sh
printf "\x1b[1;32m *\x1b[0;m [initramfs] Loading, please wait..."
export PATH=/sbin:/usr/sbin:/bin:/usr/bin
[ -d /dev ]  || mkdir -m 0755 /dev
[ -d /root ] || mkdir -m 0700 /root
[ -d /sys ]  || mkdir /sys
[ -d /proc ] || mkdir /proc
[ -d /tmp ]  || mkdir /tmp
mkdir -p /var/lock
#mount -a
mount -t sysfs -o nodev,noexec,nosuid sysfs /sys
mount -t proc -o nodev,noexec,nosuid proc /proc
/sbin/mdev -s
#clear
printf "\x1b[1;32m *\x1b[0;32m Welcome to MiniSys on Initramfs !\x1b[m\n"
exec /sbin/init
\end{verbatim}\end{framed}
With this initramfs init, you can only stay in initramfs after boot.\newline
Creating initramfs init C program:\newline
\texttt{bsdbar.h :}
\input{bsdbar.h.tex}
\texttt{init.c :}
\input{init.c.tex}
To compile it, just type
\begin{quote}
    \texttt{\$ gcc -O2 -o init init.c}
\end{quote}

\subsection{Init Script}
Use \textbf{either} inittab or rcS for busybox init here, and don't use both them.\newline
Lookup busybox init for reason\cite{bib:busybox.net}.

\subsubsection{etc/init.d/rcS}
Example for busybox init :
\begin{framed}\begin{verbatim}
#!/bin/sh
printf "\x1b[1;32m*\x1b[0;m [init] Loading, please wait..."
export PATH=/sbin:/usr/sbin:/bin:/usr/bin
mount -a
#clear
printf "\x1b[32m* Welcome to MiniSys on Initramfs !\x1b[m\n"
/bin/sh
\end{verbatim}\end{framed}

\subsubsection{/etc/inittab}
\begin{verbatim}
tty1::respawn:/sbin/getty 38400 tty1
tty2::respawn:/sbin/getty 38400 tty2
tty3::respawn:/sbin/getty 38400 tty3
tty4::respawn:/sbin/getty 38400 tty4
tty5::respawn:/sbin/getty 38400 tty5
tty6::respawn:/sbin/getty 38400 tty6
\end{verbatim}

\subsection{Wrap Initrd}
\begin{framed}\begin{verbatim}
# cd initrd/
# find . | cpio -o -H newc > ../initrd.img
# gzip -k ../initrd.img
\end{verbatim}\end{framed}
you can also gzip the image to initrd.img.gz, kernel recogonizes it too.

\section{Install the System into a USB stick}
\subsection{Partition USB stick}
Assume I have an 8GB USB Flash stick, detected as /dev/sdc.
\begin{framed}\begin{verbatim}
# parted /dev/sdc
  > mktable  (gpt)
  > mkpart (2MB-4MB as BIOS_GRUB)
  > set 1 bios_grub
  > mkpart (4MB-REST as /)
  > quit
# partprobe
# lsblk || cat /proc/partitions
\end{verbatim}\end{framed}
\subsection{Make Filesystem}
\begin{framed}\begin{verbatim}
# mkfs.ext4 /dev/sdc2 || mke2fs -t ext4 /dev/sdc2
# mount -t ext4 /dev/sdc2 /mnt
\end{verbatim}\end{framed}
\subsection{Copy Kernel and Initrd}
\begin{framed}\begin{verbatim}
# cp bzImage /mnt/boot
# cp initrd.img /mnt/boot
\end{verbatim}\end{framed}
\subsection{Install Grub on USB Stick}
\begin{framed}\begin{verbatim}
# grub-install --boot-directory /mnt/boot /dev/sdc
\end{verbatim}\end{framed}

\section{Boot Test}
\subsection{Boot via QEMU, without USB}
Test bzImage + initrd.img. 
\begin{framed}\begin{verbatim}
# qemu-system-x86_64 -enable-kvm -m 512 -kernel bzImage -initrd initrd.img
\end{verbatim}\end{framed}
\subsection{Boot via QEMU, with USB}
Test Grub2 + bzImage + initrd.img.
\begin{framed}\begin{verbatim}
# qemu-system-x86_64 -enable-kvm -m 512 -hda /dev/sdc
\end{verbatim}\end{framed}
\subsubsection{Talk with Grub2}
\begin{framed}\begin{verbatim}
grub> ls
grub> insmod linux
grub> prefix=(hd0,gpt2)/root/grub
grub> root=(hd0,gpt2)
grub> linux /boot/bzImage [OPTIONS]
grub> initrd /boot/initrd.img
grub> boot
\end{verbatim}\end{framed}
Where OPTIONS depends on your preference.
\subsubsection{Put Grub2 config into boot/grub/grub.cfg}
\begin{framed}\begin{verbatim}
# grub.cfg
insmod part_gpt
insmod ext2
set root=(hd0,gpt2)

echo "* [grub] Loading linux ...\n"
linux /boot/bzImage root=/dev/ram0 init=/sbin/init
echo "* [grub] Loading initrd.img ...\n"
initrd /boot/initrd.img
echo "* [grub] Booting ...\n"
boot
\end{verbatim}\end{framed}
Then the system would autostart as grub2 found grub.cfg.

\section{Extend the Mini System}
\subsection{Script /init in initrd.img}
Imitating Debian's script from update-initramfs and the script from Linux from scratch\cite{bib:lfs}.\newline
Note that, this script defined a new function "choose if you want to switch root", add corresponding
kernel parameter then you can activate this function:
\begin{itemize}
\item \texttt{switch} is default, means that if a root filesystem is detected, then init would switch root into it.
\item \texttt{noswitch} meas that, don't switch root even if an available root is detected.
\end{itemize}
\input{init.sh.tex}

\subsection{Prepare Stage3 rootfs}
\subsubsection{Make it myself}
"debootstrap" for Debian or Ubuntu. For example,
\begin{verbatim}
# debootstrap ubstable ./unstable-chroot http://ftp.us.debian.org/debian
\end{verbatim}

\subsubsection{Use an already cooked one}
Download the Stage3 tarball of Archlinux\footnote{https://www.archlinux.org/} 
or Gentoo\footnote{http://www.gentoo.org/}.\newline
Here we can use both of them. And this is hint:
\begin{itemize}
\item \textbf{Archlinux} stage3 does not symlink /usr/lib/systemd/systemd to /sbin/init,
	so you may encounter a kernel panic if you don't modify my initramfs init script.
	to avoid this just set the init parameter.
\item \textbf{Gentoo} stage3 tarballs works well.
\item \textbf{Debian} stage3 tarballs I made also works well.
\end{itemize}
\subsubsection{Make the disk image OR copy them into USB}
You can extrace the Stage3 tarball into a disk image:
\begin{framed}\begin{verbatim}
# dd of=disk.img bs=1 seek=4G count=0
# mkfs.ext4 disk.img
# mount disk.img /mnt
# tar zxvf Stage3.tar.gz -C /mnt
# do some configurations
# umount /mnt
\end{verbatim}\end{framed}
or just extract it into you USB stick.

\subsection{QEMU: Boot with the new disk}
\begin{verbatim}
# qemu-system-x86_64 -enable-kvm -m 512 -kernel bzImage -initrd initrd.img
    -hda disk.img -append "root=/dev/sda init=/usr/lib/systemd/systemd"
\end{verbatim}

\subsection{QEMU: Boot from the USB}
\begin{verbatim}
# qemu-system-x86_64 -enable-kvm -m 512
    -hda disk.img -append "root=/dev/sda init=/usr/lib/systemd/systemd"
\end{verbatim}

\section{Compile Static Bash}
\subsection{Download bash source}
bash-4.3.tar.gz from GNU.
\subsection{Compile static Bash}
\begin{verbatim}
$ ./configure --enable-static-link --without-bash-malloc
\end{verbatim}
Then you will see bash as following:
\begin{verbatim}
$ ldd bash
bash: ELF 64-bit LSB executable, x86-64, version 1 (GNU/Linux), 
      statically linked, for GNU/Linux 2.6.32, 
      BuildID[sha1]=ab5bcc419a27e6c54d0fb352c28019446e68dd46, not stripped
\end{verbatim}
\subsection{Make bash.tar.gz tarball}
I suggest copy those files into directory bash.pkg:
\begin{itemize}
\item staticly linked bash excutable
\item examples/startup-files/Bash\_profile
\item examples/startup-files/bashrc
\end{itemize}
then make dir bash.pkg as bash.tar.gz
\subsection{Install static bash into initrd}
Just extract the tarball into initrd/.

\section{Lazy Glibc supporting Sed}
If, say, I want to use the program GNU Sed in the freshly cooked system,
but we lack the glibc that supporting sed. so we can use Debian's precompiled glibc,
and a sed that customed by us (deleting .so links that we don't like.).
\subsection{Compiling GNU Sed}
\begin{verbatim}
$ ./configure --without-selinux --disable-acl
$ make -j4
$ ldd sed/sed
  linux-vdso.so.1 (0x00007ffcec738000)
  libc.so.6 => /lib/x86_64-linux-gnu/libc.so.6 (0x00007f870f03b000)
  /lib64/ld-linux-x86-64.so.2 (0x00007f870f410000)
$ mkdir sed.pkg/
\end{verbatim}
See, it's ldd output is very clean. Now let's add the libc.\newline
Then you should copy files you need into dir sed.pkg/

\subsection{Wrap GLibc}
Now let's grab a libc to support the sed (and other programs depending on this libc).\newline
For convenience, I just downloaded the Debian's precompiled glibc, and re-packed the glibc together with sed.
\begin{verbatim}
$ apt-get download libc6 libc-bin
$ for DEB in $(ls *.deb); do dpkg -X ${DEB} sed.pkg/; done
$ cd sed.pkg
$ tar zcvf ../sed.tar.gz .
\end{verbatim}

\subsection{Install libc + sed into initrd}
Just extract the tarball into initrd/

% ---------------------------------------------------------------------------------
\appendix
\newpage
\section{Base64 Code of /init}
Following is init.base64 :

% insert the code file
\input{init.base64.tex}

To decode it, type :
\begin{verbatim}
$ base64 -d init.base64
\end{verbatim}

\newpage
\section{Misc}
\begin{thebibliography}{10}
	\bibitem{bib:kernel.org} The Linux Kernel Archive\\https://kernel.org
	\bibitem{bib:busybox.net} Busybox\\http://www.busybox.net
	\bibitem{bib:kernel.doc} Linux kernel Doc\\https://kernel.org/doc\\You can also lookup the Documentation dir in kernel source tree.
	\bibitem{bib:kernel.doc.boot} Linux kernel boot protocol\\linux-4.0/Documentation/./x86/boot.txt
	\bibitem{bib:grub.doc} GNU : grub document\\http://www.gnu.org/software/grub/manual/grub.html
	\bibitem{bib:lfs} LFS : Linux From Scratch\\http://www.linuxfromscratch.org/lfs/
	\bibitem{bib:blfs} BLFS : Beyond Linux From Scratch\\http://www.linuxfromscratch.org/blfs/
	\bibitem{ } A Guide on initramfs from LFS\\http://www.linuxfromscratch.org/hints/downloads/files/initramfs.txt
	\bibitem{ } A Guide on initrd from LFS\\http://www.linuxfromscratch.org/hints/downloads/files/initrd.txt
	\bibitem{bib:gnu.bash} Download GNU Bash\\http://ftp.gnu.org/gnu/bash/
	\bibitem{ } The kernel module dir\\/lib/modules/\$(uname -r)/
	\bibitem{ } An Article about Grub\\http://blog.csdn.net/guanggY/article/details/6210774
	\bibitem{ } An Article about Initramfs\\http://blog.csdn.net/lvqqrainbow/article/details/6536422
	\bibitem{ } An Article about Kernel Boot\\http://blog.csdn.net/zhoudaxia/article/details/6666683
	\bibitem{ } An Article about Kernel Boot\\http://blog.sina.com.cn/s/blog\_b02f77c80101db1t.html
\end{thebibliography}

\section{LISENCE}
\begin{verbatim}
The MIT LISENCE.
Copyright (C) 2015 Lumin Zhou
\end{verbatim}

\end{document}
% ----end document----
